---
title: "Olist_Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---


```{r setup, include=FALSE}
library(flexdashboard)
```

# EDA on Customer Orders

## Column {.tabset}

### Trend of number of users per month per year

```{r}
chartA
```

### Revenue by Month by Year

```{r}
hc_revenue
```

### Popular products

```{r}
hc_prod
```



## Column {data-width = 250}

### Type of payments
```{r}
hc_pay
```

### State with highest number of orders
```{r}
state
```

# EDA on Customer Reviews

## Column { .tabset}


### Reviews and lengths
```{r}
hc_score
```

### Distribution of review_scores

```{r}
bar_reviews
```

## Column {data-width = 250}

### % of review_score as 1
```{r}
gauge(12, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
))
```

### % of review_score as 5

```{r}
gauge(57, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
))
```

### Average review_score

```{r}
valueBox(4.07, icon = "fa-pencil", href="#details")
```



# Seller Side EDA

## Column {data-width = 500}

### MQL by different channels
```{r}
ch_ac
```

## Column {data-width = 500}

### MQL volume
```{r}
mq_vol
```



### Closed_deals volume
```{r}
cd
```

## Column {data-width = 500}

### Type of sellers
```{r}
sel_type
```


# Forecasting

## Column {.tabset}

### Time_Series plot
```{r}
plot(mrev_olist_ts)
```



### Trend cycle

```{r}
tc
```

### Decomposition

```{r}
rev_dc
```


### ACF & PACF

```{r}
rev_olist %>% gg_tsdisplay(revenue, plot_type = 'partial')
```


## Column {.tabset}

### Differenced Time_Series

```{r}
rev_olist %>% mutate(diff1 = difference(revenue),lag=12) %>%
  mutate(diff2 = difference(diff1))%>%
  gg_tsdisplay(diff2, plot_type = 'partial')
```


### ARIMA forecast
```{r}
fore
```

### ETS(M,A,N) forecast
```{r}
ets_fore
```





# Customer_Segmentation


## Column {data-width = 250}

### Maximum Recency
```{r}
valueBox(766, href="#details")
```

### Minimum Recency

```{r}
valueBox(17, href="#details")
```

### Maximum Frequency

```{r}
valueBox(16,href="#details",icon = "fa-line-chart",color = "info")
```

### Minimum Frequency

```{r}
valueBox(1,href="#details", icon = "fa-line-chart",color = "info")
```

### Maximum Monetary_Value

```{r}
valueBox(109321,href="#details",icon = "fa-money", color ='green')
```

### Minimum Monetary_Value


```{r}
valueBox(9.59, href="#details",icon = "fa-money", color ='green')
```

## Column {.tabset}

### Recency based Customer Segmentation
```{r}
rec_seg
```



### K means based Customer Segmentation

```{r}
final_seg
```

